---
title: "Lab 10 Homework"
author: "Min-Yao Jhu"
date: "`r Sys.Date()`"
output:
  html_document: 
    theme: spacelab
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instructions
Answer the following questions and complete the exercises in RMarkdown. Please embed all of your code and push your final work to your repository. Your final lab report should be organized, clean, and run free from errors. Remember, you must remove the `#` for the included code chunks to run. Be sure to add your name to the author header above. For any included plots, make sure they are clearly labeled. You are free to use any plot type that you feel best communicates the results of your analysis.  

Make sure to use the formatting conventions of RMarkdown to make your report neat and clean!  

## Load the libraries
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)
library(here)
library(naniar)
``` 

## Desert Ecology
For this assignment, we are going to use a modified data set on [desert ecology](http://esapubs.org/archive/ecol/E090/118/). The data are from: S. K. Morgan Ernest, Thomas J. Valone, and James H. Brown. 2009. Long-term monitoring and experimental manipulation of a Chihuahuan Desert ecosystem near Portal, Arizona, USA. Ecology 90:1708.
```{r}
deserts <- read_csv(here("lab10", "data", "surveys_complete.csv"))
```

1. Use the function(s) of your choice to get an idea of its structure, including how NA's are treated. Are the data tidy?  

```{r}
str(deserts)
```
```{r}
skimr::skim(deserts)
```
```{r}
naniar::miss_var_summary(deserts)
```
```{r}
summary(deserts)
```



2. How many genera and species are represented in the data? What are the total number of observations? Which species is most/ least frequently sampled in the study?

```{r}
deserts %>% 
  summarise(n_genera = n_distinct(genus),
            n_species = n_distinct(species),
            n = n())
```
```{r}
deserts %>% 
  tabyl(species) %>% 
  as.data.frame() %>% 
  arrange(desc(n))
```


3. What is the proportion of taxa included in this study? Show a table and plot that reflects this count.

```{r}
deserts %>% 
  tabyl(taxa) %>% 
  as.data.frame()
```

```{r}
deserts %>% 
  ggplot(aes(x = taxa, fill = taxa)) + 
  geom_bar() +
  scale_y_log10() +
  labs(title = "Observations by taxa in Deserts Data",
       x = "island name",
       y = "log10 count")
```

4. For the taxa included in the study, use the fill option to show the proportion of individuals sampled by `plot_type.`

```{r}
deserts %>% 
  ggplot(aes(x = taxa, fill = plot_type)) + 
  geom_bar() +
  scale_y_log10() +
  labs(title = "Observations by taxa in Deserts Data",
       x = "taxa",
       y = "log10 count")
```
```{r}
deserts %>% 
  ggplot(aes(x = taxa, fill = plot_type)) + 
  geom_bar(position = "dodge") +
  scale_y_log10() +
  labs(title = "Observations by taxa in Deserts Data",
       x = "taxa",
       y = "log10 count")
```

5. What is the range of weight for each species included in the study? Remove any observations of weight that are NA so they do not show up in the plot.

```{r}
deserts %>%
  filter(weight!="NA") %>%
  group_by(species) %>% 
  summarise(mean_weight = mean(weight, na.rm = T),
            min_weight = min(weight, na.rm = T),
            max_weight = max(weight, na.rm = T),
            median_weight = median(weight, na.rm = T))
```
```{r}
deserts %>%
  filter(weight!="NA") %>%
  ggplot(aes(x = species, y = weight, fill = species)) + 
  geom_boxplot() +
  coord_flip()+  
  labs(title = "Weight range by species in Deserts Data",
       x = "species name",
       y = "weight")
```

```{r}
deserts %>%
  filter(weight!="NA") %>%
  ggplot(aes(x = species, y = weight, fill = species)) + 
  geom_boxplot() +
  coord_flip() +
  scale_y_log10() +
  labs(title = "Weight range by species in Deserts Data",
       x = "species name",
       y = "log10 weight")
```

6. Add another layer to your answer from #5 using `geom_point` to get an idea of how many measurements were taken for each species.

```{r}
deserts %>%
  filter(weight!="NA") %>%
  ggplot(aes(x = species, y = weight)) + 
  geom_boxplot() +
  geom_point(alpha=0.01) +
  coord_flip()+  
  labs(title = "Weight range by species in Deserts Data",
       x = "species name",
       y = "weight")
```
```{r}
deserts %>%
  filter(weight!="NA") %>%
  ggplot(aes(x = species, y = weight)) + 
  geom_boxplot() +
  geom_jitter(alpha=0.01) +
  coord_flip()+  
  labs(title = "Weight range by species in Deserts Data",
       x = "species name",
       y = "weight")
```
```{r}
deserts %>%
  filter(weight!="NA") %>%
  ggplot(aes(x = species, y = weight)) + 
  geom_boxplot() +
  geom_jitter(width = 0.1, height = 0.1, alpha=0.01) +
  coord_flip()+  
  labs(title = "Weight range by species in Deserts Data",
       x = "species name",
       y = "weight")
```

```{r}
deserts %>%
  filter(weight!="NA") %>%
  ggplot(aes(x = species, y = weight)) + 
  geom_boxplot() +
  geom_point() +
  coord_flip() +
  scale_y_log10() +
  labs(title = "Weight range by species in Deserts Data",
       x = "species name",
       y = "log10 weight")
```

7. [Dipodomys merriami](https://en.wikipedia.org/wiki/Merriam's_kangaroo_rat) is the most frequently sampled animal in the study. How have the number of observations of this species changed over the years included in the study?

```{r}
deserts %>%
  filter(species=="merriami") %>% 
  group_by(year) %>% 
  summarise(n=n())
```
```{r}
deserts %>%
  filter(species=="merriami") %>% 
  ggplot(aes(x = year)) + 
  geom_bar() +
  labs(title = "Observations by year in Deserts Data",
       x = "year",
       y = "count")
```

```{r}
deserts %>%
  filter(species=="merriami") %>% 
  group_by(year) %>% 
  summarise(n=n()) %>% 
  ggplot(aes(x = year,y=n)) + 
  geom_col() +
  labs(title = "Observations by year in Deserts Data",
       x = "year",
       y = "count")
```

8. What is the relationship between `weight` and `hindfoot` length? Consider whether or not over plotting is an issue.

> Overplotting is when the data or labels in a data visualization overlap, making it difficult to see individual data points in a data visualization. Overplotting typically occurs when there are either a large number of data points and/or a small number of unique values in the dataset.

> Fixes for overplotting include reducing the size of points, changing the shape of points, jittering, tiling, making points transparent, only showing a subset of points, and using algorithms to prevent labels from overlapping.

```{r}
deserts %>%
  ggplot(aes(x=weight, y=hindfoot_length))+
  geom_point(na.rm=T)+
  geom_smooth(method = "lm",na.rm=T) +
  labs(title = "weight vs. hindfoot in Deserts Data",
       x = "year",
       y = "count")
```
> Too many data points with similar values
> Data points with a small number of unique values
> Overlapping labels

> solution1: Reduce the size of points

```{r}
deserts %>%
  ggplot(aes(x=weight, y=hindfoot_length))+
  geom_point(size = 0.8, na.rm=T)+
  geom_smooth(method = "lm",na.rm=T)
```

> Showing a subset of the data points
> Changing plotting symbols (glyphs)
> Use transparency to deal with overplotting

```{r}
deserts %>%
  ggplot(aes(x=weight, y=hindfoot_length))+
  geom_point(na.rm=T, alpha=0.1)+
  geom_smooth(method = "lm",na.rm=T)
```

> Jittering

```{r}
deserts %>%
  ggplot(aes(x=weight, y=hindfoot_length))+
  geom_jitter(na.rm=T, alpha=0.1)+
  geom_smooth(method = "lm",na.rm=T)
```


9. Which two species have, on average, the highest weight? Once you have identified them, make a new column that is a ratio of `weight` to `hindfoot_length`. Make a plot that shows the range of this new ratio and fill by sex.

```{r}
deserts %>%
  filter(weight!="NA") %>%
  group_by(species) %>% 
  summarise(mean_weight = mean(weight, na.rm = T)) %>% 
  arrange(desc(mean_weight))
```
> albigula and spectabilis

```{r}
deserts2 <- deserts %>%
  filter(weight!="NA" & hindfoot_length!="NA") %>%
  mutate(w_h_ratio = weight/hindfoot_length)
deserts2
```

```{r}
deserts2 %>% 
  filter(species=="albigula"|species=="spectabilis") %>% 
  ggplot(aes(x = species, y = w_h_ratio, fill = sex)) + 
  geom_boxplot() +
  labs(title = "Ratio of weight to hindfoot_length by species in Deserts Data",
       x = "species name",
       y = "Ratio (weight/hindfoot_length)")
```
```{r}
deserts2 %>% 
  filter(species=="albigula"|species=="spectabilis") %>% 
  filter(sex!="NA") %>% 
  ggplot(aes(x = species, y = w_h_ratio, fill = sex)) + 
  geom_boxplot() +
  labs(title = "Ratio of weight to hindfoot_length by species in Deserts Data",
       x = "species name",
       y = "Ratio (weight/hindfoot_length)")
```
```{r}
deserts2 %>% 
  filter(species=="albigula"|species=="spectabilis") %>% 
  filter(sex!="NA") %>% 
  ggplot(aes(x = species, y = w_h_ratio, color = sex, shape = sex)) + 
  geom_jitter() +
  labs(title = "Ratio of weight to hindfoot_length by species in Deserts Data",
       x = "species name",
       y = "Ratio (weight/hindfoot_length)")
```

10. Make one plot of your choice! Make sure to include at least two of the aesthetics options you have learned.


## Push your final code to GitHub!
Please be sure that you check the `keep md` file in the knit preferences. 