---
title: "loop"
author: "Min-Yao"
date: "2021/1/19"
output: 
  html_document: 
    keep_md: yes
---

## Learning Goals
*At the end of this exercise, you will be able to:*  
1. know the concept of loops.
2. understand the three different types of loops
3. use loops in vectors, matrixes, and data frames.
4. use loops for plotting.
5. use loops to combine multiple files.

## Resources and references

- [Loops in R](https://statisticsglobe.com/loops-in-r/)

## Load the libraries
```{r message=FALSE, warning=FALSE}
library("tidyverse")
```

## What Are Loops?
A loop is a programming instruction that repeats until a specific condition is reached.
The loop executes a code block again and again until no further action is required.
Each time the code block within the loop is executed is called an iteration.

## Different Types of Loops
Depending on your specific programming situation, you may need different loop-structures that execute the code blocks within the loop on the basis of different conditions.

The R programming language generally provides three different types of loops: 
1. for-loops
2. while-loops
3. repeat-loops

## For-Loops

for-loops specify a collection of objects (e.g. elements in a vector or list) to which a code block should be applied.

## Syntax of for loop

for (variable in vector) {
  do something
}

Using the names above, each iteration of variable takes the value of one of the elements of vector. Here’s one example:

#### example 0: Loop Through Vector in R (Basics)

```{r}
for (i in 1:10) {
  print(i^2)
}
```

#### example 1. count the number of even numbers in a vector

```{r}
x <- c(1,2,5,3,10,9,8,11,6)
count <- 0

for (i in x) {
  if(i %% 2 == 0)  
    print(i)
}
```

```{r}
for (i in x) {
  if(i %% 2 == 0)  
    count = count+1
}

print(count)
```
```{r}
random_number <- sample(1:1000, size = 100)
random_number

for (i in random_number) {
  if(i %% 2 == 0)  
    count = count+1
}

print(count)
```
### for-Loop Through Columns of Data Frame

```{r}
data <- data.frame(x1 = 1:5,    # Create example data
                   x2 = 6:10,
                   x3 = 11:15)
data 
```

```{r}
data1 <- data # Replicate example data

for(i in 1:ncol(data1)) {       # for-loop over columns
  data1[ , i] <- data1[ , i] + 10
}
data1
```

### for-Loop Over Rows of Data Frame

```{r}
data2 <- data                   # Replicate example data

for(i in 1:nrow(data2)) {       # for-loop over rows
  data2[i, ] <- data2[i, ] - 100
}
data2
```

### for-Loop Over Data Frame Columns

```{r}
data(iris)                                                 # Loading iris flower data set
head(iris)                                                 # Inspecting iris flower data set
```

```{r}
iris1 <- iris                                          # Replicate iris data set

for(i in 1:ncol(iris1)) {                              # Head of for-loop
 
  if(grepl("Width", colnames(iris1)[i])) {             # Logical condition
 
    iris1[ , i] <- iris1[ , i] * 10              # Code block
  }
}

head(iris1)
```

### Example 9: Creating Multiple Plots within for-Loop

```{r}
for(i in 1:(ncol(iris) - 1)) {                             # Head of for-loop
 
  plot(1:nrow(iris), iris[ , i])                           # Code block
  Sys.sleep(1)                                             # Pause code execution
}
```


### Looping over files

#### Let’s download some simulated satellite collar data

```{r}
#download.file("http://www.datacarpentry.org/semester-biology/data/locations.zip", "locations.zip")
#unzip("locations.zip")
```

```{r}
list.files()
```

```{r}
data_files <- list.files(pattern = "locations-2*", 
                        full.names = TRUE)
data_files
str(data_files)
```
```{r}
results <- vector(mode = "integer", length = length(data_files))
results
```

```{r}
for (i in 1:length(data_files)){
  data0 <- read_csv(data_files[i])
  count <- nrow(data0)
  results[i] <- count
}
results
```

https://datacarpentry.org/semester-biology/materials/for-loops-R/

https://datacarpentry.org/semester-biology/exercises/Loops-multi-file-analysis-R/

### Storing loop results in a data frame

```{r}
results <- data.frame(file_name = vector(mode = "character", length = length(data_files)),
                      count = vector(mode = "integer", length = length(data_files)),
                      long = vector(mode = "numeric", length = length(data_files)))
results
```

```{r}
for (i in 1:length(data_files)){
  data0 <- read_csv(data_files[i])
  count <- nrow(data0)
  results$file_name[i] <- data_files[i]
  results$count[i] <- count
  results$long[i] <- data0$long
  results$lat[i] <- data0$lat
}
results
```

### loops practice: Multi-file Analysis

1. You have a satellite collars on a number of different individuals and want to be able to quickly look at all of their recent movements at once. The data is posted daily to a url as a zip file that contains one csv file for each individual: http://www.datacarpentry.org/semester-biology/data/individual_collar_data.zip Start your solution by:

* Downloading the zip file using download.file()
* Unziping it using unzip()
* Obtaining a list of all of the files with file names matching the pattern "collar-data-.*.txt"

```{r}
#download.file("http://www.datacarpentry.org/semester-biology/data/individual_collar_data.zip", "collar.zip")
#unzip("collar.zip")
```
2. Use a loop to load each of these files into R and make a line plot (using geom_path()) for each file with long on the x axis and lat on the y axis. Graphs, like other types of output, won’t display inside a loop unless you explicitly display them, so you need put your ggplot() command inside a print() statement. Include the name of the file in the graph as the graph title using labs().

```{r}

```


## While-Loops

while-loops repeat a code block as long as a certain logical condition is TRUE.

This code is typically used when we don’t know the exact number of times our R code needs to be executed.

### example 0: Writing while-Loop in R (Basics)

```{r}
x_while <- 0                  # Preliminary specification of data object

while(x_while < 10) {         # Head of while-loop
 
  x_while <- x_while + 1      # Body of while-loop
  print(x_while)
}

```
### example 1: while-Loop Through Columns of Data Frame

```{r}
data3 <- data                   # Replicate example data

i <- 1                          # Create running index

while(i <= 2) {                  # Start while-loop
  data3[ , i] <- data3[ , i] + 100
  i <- i + 1
}

data3
```

### example 2: Running while-Loop Through Data Frame Columns

```{r}
data(iris)                                        # Loading exemplifying data set
head(iris)                                        # Showing head of data in RStudio console
iris2 <- iris                   # Replicate example data
```

```{r}
running_index <- 1                                       # Create index

while(is.numeric(iris2[ , running_index])) {              # Start of while-loop
 
  iris2[ , running_index] <- iris2[ , running_index] * 10  # Code block
  running_index <- running_index + 1
 
}

head(iris2)                                               # Showing head of updated data
```



## Repeat-Loops

repeat-loops repeat a code block until a break condition is fulfilled. This break condition marks the end of the loop.

repeat-loops follow a similar logic as while-loops, since they can also be used when the user doesn’t know the exact number of times the R code should be repeated. However, repeat-loops are not as popular as while-loops.

### example 0: Writing a repeat-Loop in R (Basics)

```{r}
x_repeat <- 0                 # Preliminary specification of data object

repeat{                       # Head of repeat-loop
 
  x_repeat <- x_repeat + 1    # Body of repeat-loop
  print(x_repeat)
 
  if(x_repeat >= 10) {        # Break condition of repeat-loop
 
    break
  }
}
```
### example 1: repeat-Loop Through Columns of Data Frame

```{r}
data4 <- data                   # Replicate example data

i <- 1                          # Create running index

repeat{                         # Start repeat-loop
 
  data4[ , i] <- data4[ , i] + 100
  i <- i + 1
 
  if(i > 2) {
    
    break
  }
}

data4
```

### example 2: Running repeat-Loop Over Data Frame Columns

```{r}
data(iris)                          # Loading exemplifying data frame
head(iris)                          # Head of example data

iris3 <- iris                   # Replicate example data
```

```{r}
i <- 0                              # Create running index

repeat{                             # Start
 
  i <- i + 1                        # Update running index
 
  if(!is.numeric(iris3[ , i])) {     # Break condition
 
    break
  }
 
  iris3[ , i] <- iris3[ , i] * 10    # Change data
}
head(iris3)
```


### other practice questions

```{r}
library(tidyverse)
```

## Load the data
For these exercises we will use the data from lab 4_1.  
```{r}
fish <- readr::read_csv("data/Gaeta_etal_CLC_data.csv")
mammals <- readr::read_csv("data/mammal_lifehistories_v2.csv")
```

To keep things tidy, I am going to rename the variables in the mammals data.  
```{r}
mammals <- rename(mammals, genus=Genus, wean_mass="wean mass", max_life="max. life", litter_size="litter size", litters_per_year="litters/year")
```

```{r}
str(mammals)
str(fish)
```

## That's it! Take a break and I will see you on Zoom!  

-->[Home](https://jmledford3115.github.io/datascibiol/)
