---
title: "loop"
author: "Min-Yao"
date: "2021/1/19"
output: html_document
---

# A Tutorial on Loops in R

What Are Loops?

“Looping”, “cycling”, “iterating” or just replicating instructions is an old practice that originated well before the invention of computers. It is nothing more than automating a multi-step process by organizing sequences of actions or ‘batch’ processes and by grouping the parts that need to be repeated.


## Different Types of Loops

Depending on your specific programming situation, you may need different loop-structures that execute the code blocks within the loop on the basis of different conditions.

The R programming language generally provides three different types of loops: for-loops, while-loops, and repeat-loops. 

## For-Loops

### Syntax of for loop

```{r}
#for (variable in vector) {
#  do something
#}
```

> Using the names above, each iteration of variable takes the value of one of the elements of vector. Here’s one example:

#### example 0.

```{r}
for (i in 1:10) {
  print(i^2)
}
```

#### example 1. count the number of even numbers in a vector

```{r}
x <- c(1,2,5,3,10,9,8,11,6)
count <- 0

for (i in x) {
  if(i %% 2 == 0)  
    print(i)
}
```

```{r}
for (i in x) {
  if(i %% 2 == 0)  
    count = count+1
}

print(count)
```
```{r}
random_number <- sample(1:1000, size = 100)
random_number

for (i in random_number) {
  if(i %% 2 == 0)  
    count = count+1
}

print(count)
```

```{r}

```


## While-Loops

```{r}

```


## Repeat-Loops

```{r}

```

```{r}
library(tidyverse)
```

## Load the data
For these exercises we will use the data from lab 4_1.  
```{r}
fish <- readr::read_csv("data/Gaeta_etal_CLC_data.csv")
mammals <- readr::read_csv("data/mammal_lifehistories_v2.csv")
```

To keep things tidy, I am going to rename the variables in the mammals data.  
```{r}
mammals <- rename(mammals, genus=Genus, wean_mass="wean mass", max_life="max. life", litter_size="litter size", litters_per_year="litters/year")
```

```{r}
str(mammals)
str(fish)
```
