---
title: "genetic_lab"
author: "Min-Yao"
date: "2021/2/17"
output: html_document
---

In practice, one would use the function read.cross() to load a data set into R. See the help
ﬁle (type ?read.cross), look at the sample data sets at https://rqtl.org/sampledata,
and consider Chapter 2 of Broman and Sen (2009) A Guide to QTL Mapping with R/qtl
(https://rqtl.org/book).

A data ﬁle for the mapthis data is at https://rqtl.org/tutorials/mapthis.csv.

```{r}
library(tidyverse)
library(ggplot2)
library(qtl)
library(DataExplorer)
data(mapthis)
```

```{r}
str(mapthis)
glimpse(mapthis)
```

```{r}
summary(mapthis)
```

```{r}
plotMissing(mapthis)
```

```{r}
plot_missing(mapthis, missing_only = TRUE)
```


```{r}
ntyped(mapthis)
```

The function ntyped() provides the numbers of genotyped markers for each individual (or
the number of genotyped individuals for each marker).

```{r}
plot(ntyped(mapthis), ylab="No. typed markers", main="No. genotypes by individual")
```

```{r}
ntyped(mapthis) %>% 
  as.data.frame() %>% 
  rename(genotyped_marker_number=".") %>% 
  rowid_to_column(var = "individual_number") %>% 
  ggplot(aes(x=individual_number,y=genotyped_marker_number))+
  geom_point()
```

```{r}
plot(ntyped(mapthis, "mar"), ylab="No. typed individuals", main="No. genotypes by marker")
```

```{r}
ntyped(mapthis, "mar")
```

```{r}
cg <- comparegeno(mapthis)
hist(cg[lower.tri(cg)], breaks=seq(0, 1, len=101), xlab="No. matching genotypes")
rug(cg[lower.tri(cg)])
```

